```{r load-data}
data <- read.csv("~/Downloads/UniversalBank.csv")
head(data)
```
```{r clean-data}
clean_data <- subset(data, select = -c(ID, ZIP.Code))
clean_data$Personal.Loan <- as.factor(clean_data$Personal.Loan)
str(clean_data)
```
```{r split-data}
set.seed(123)
library(caret)
train_index <- createDataPartition(clean_data$Personal.Loan, p = 0.7, list = FALSE)
train_data <- clean_data[train_index, ]
test_data <- clean_data[-train_index, ]
```
```{r knn-model}
library(class)
normalize <- function(x) {
  return ((x - min(x)) / (max(x) - min(x)))
}

train_X <- as.data.frame(lapply(train_data[, -which(names(train_data) == "Personal.Loan")], normalize))
test_X  <- as.data.frame(lapply(test_data[, -which(names(test_data) == "Personal.Loan")], normalize))

train_Y <- train_data$Personal.Loan
test_Y  <- test_data$Personal.Loan

library(class)
knn_pred <- knn(train = train_X, test = test_X, cl = train_Y, k = 5)

table(Predicted = knn_pred, Actual = test_Y)
conf_matrix <- table(Predicted = knn_pred, Actual = test_Y)

accuracy <- sum(diag(conf_matrix)) / sum(conf_matrix)
sensitivity <- conf_matrix["1", "1"] / sum(conf_matrix[, "1"])
specificity <- conf_matrix["0", "0"] / sum(conf_matrix[, "0"])

accuracy
sensitivity
specificity
ks <- c(1, 3, 5, 7, 15)

results <- data.frame(k = ks, Accuracy = NA, Sensitivity = NA, Specificity = NA)

for (i in seq_along(ks)) {
  k_val <- ks[i]
  knn_pred <- knn(train = train_X, test = test_X, cl = train_Y, k = k_val)
  conf_matrix <- table(Predicted = knn_pred, Actual = test_Y)
  
  acc <- sum(diag(conf_matrix)) / sum(conf_matrix)
  sens <- conf_matrix["1", "1"] / sum(conf_matrix[, "1"])
  spec <- conf_matrix["0", "0"] / sum(conf_matrix[, "0"])
  
  results[i, ] <- c(k_val, acc, sens, spec)
}

results
misclassified <- test_data[knn_pred != test_Y, ]
nrow(misclassified)
head(misclassified)
summary(misclassified)
prop.table(table(misclassified$Personal.Loan))
library(ggplot2)

ggplot(results, aes(x = k, y = Accuracy)) +
  geom_line() +
  geom_point() +
  ylim(0.9, 1) +
  labs(title = "Accuracy vs. k", x = "k", y = "Accuracy")
set.seed(123)
train_index <- createDataPartition(clean_data$Personal.Loan, p = 0.5, list = FALSE)
train_data <- clean_data[train_index, ]
remaining_data <- clean_data[-train_index, ]

val_index <- createDataPartition(remaining_data$Personal.Loan, p = 0.6, list = FALSE)
val_data <- remaining_data[val_index, ]
test_data <- remaining_data[-val_index, ]

normalize <- function(x) {
  return((x - min(x)) / (max(x) - min(x)))
}

train_X <- as.data.frame(lapply(train_data[, -which(names(train_data) == "Personal.Loan")], normalize))
val_X <- as.data.frame(lapply(val_data[, -which(names(val_data) == "Personal.Loan")], normalize))
test_X <- as.data.frame(lapply(test_data[, -which(names(test_data) == "Personal.Loan")], normalize))

train_Y <- train_data$Personal.Loan
val_Y <- val_data$Personal.Loan
test_Y <- test_data$Personal.Loan

knn_train <- knn(train = train_X, test = train_X, cl = train_Y, k = 3)
knn_val <- knn(train = train_X, test = val_X, cl = train_Y, k = 3)
knn_test <- knn(train = train_X, test = test_X, cl = train_Y, k = 3)

conf_train <- table(Predicted = knn_train, Actual = train_Y)
conf_val <- table(Predicted = knn_val, Actual = val_Y)
conf_test <- table(Predicted = knn_test, Actual = test_Y)

conf_train
conf_val
conf_test
